---
title: "ps_2"
author: "James Hutt"
date: "07/02/2020"
output: html_document
code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary libraries with tidyverse last to avoid clashes
library(fivethirtyeight)
library(gov.1005.data)
library(ggthemes)
library(tidyverse)
```

```{r madlib_answers, echo = FALSE}

# filter the data set for each madlib using pull or summarise to extract necessary data

x1 <- endorsements_2020 %>% 
  arrange(date) %>% 
  slice(1) %>% 
  pull(date)

x2 <- endorsements_2020 %>% 
  arrange(desc(date)) %>% 
  slice(1) %>% 
  pull(date)

x3 <- endorsements_2020 %>% 
  filter(position == "mayor", endorsee == "Pete Buttigieg") %>% 
  slice(1) %>% 
  pull(endorser)

x4 <- endorsements_2020 %>% 
  filter(state == "NH") %>% 
  summarise(mean(points))

x5 <- endorsements_2020 %>% 
  filter(endorsee %in% c("Amy Klobuchar", "Elizabeth Warren")) %>%
  summarise(sum(points)) 

# sapply function pulls out the class of the variable, having previously selected the one of interest. 

x6a <- endorsements_2020 %>% 
  select(position) %>% 
  sapply(class)

x6b<- endorsements_2020 %>% 
  select(endorser) %>% 
  sapply(class)

x6c <- endorsements_2020 %>% 
  select(points) %>% 
  sapply(class)

```

## Mad Libs

##### Fill in the answers to these Mad Libs, using the commands that we have learned in the readings and in class.

1. The earliest date for an endorsement in the dataset is `r x1`

2. The most recent date for an endorsement is `r x2`

3. `r x3` is the mayor that has endorsed Pete Buttigieg.

4. Endorsers from New Hampshire have an average of `r x4` endorsement points. 

5. Between them, the two female candidates who remain in the race (Amy Klobuchar and Elizabeth Warren) have a total of `r x5` endorsement points.

6. 
a) The variable “position” is a `r x6a` class of variable.  

b) The variable “endorser” is a `r x6b` class of variable.  

c) The variable “points” is a `r x6c` class of variable.  



## Put It Into Practice 
```{r congress_chart}

# Create the avg_age variable per party, per year

ages <- gov.1005.data::congress %>%
  filter(party %in% c("R", "D")) %>%
  mutate(party = as.factor(party)) %>%
  group_by(year, party) %>%
  summarize(avg_age = mean(age))

# Create the plot, of type line with the theme_fivethirtyeight and manipulate other aesthetics

ggplot(ages, aes(year, avg_age, color = party)) + 
  geom_line() + 
  scale_colour_manual(values = c("D" = "blue2", "R" = "red2")) + 
  labs(title = "Average Age of Members of Congress", 
       subtitle = "At start of term, 1947-2013",
       caption = "Source: fivethirtyeight package") + 
  theme_fivethirtyeight() + 
  scale_x_continuous(limits = c(1945, 2015), 
                     breaks = c(1950, 1960, 1970, 1980, 1990, 2000, 2010)) + 
  scale_y_continuous(limits = c(40, 63), 
                     breaks = c(40, 45, 50, 55, 60)) + 
  theme(legend.position = "none") + 
  annotate("text", x = 2000, y = 58, 
           label = "Democrats", color = "blue2") + 
  annotate("text", x = 2010, y = 53, 
           label = "Republicans", color = "red2") 
  



```

